---
title: "Exploration & Data Analysis"
author: "Pier Lorenzo Paracchini"
date: "11/27/2016"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = T, warning = F, message = F)
```

Using the `tidyverse` package for performing exploration and data analysis.

```{r requiredLibraries}
require(data.table)
require(lubridate)
require(dplyr)
require(knitr)
require(ggplot2)
input_file <- "./../../data/raw/train_ver2.csv"

#Splitting the training data as one file till 2016.04 (included)
input_train <- "./../../data/processed/df_train.RData"
input_train_typesSet <- "./../../data/processed/df_train_typesSet.RData"

#Splitting up the data by periods to have smaller chunks
#2015 split in qurters
#2016 from Jan to April (included)
input_2015_Q1 <- "./../../data/processed/df_2015_Q1.RData"
input_2015_Q2 <- "./../../data/processed/df_2015_Q2.RData"
input_2015_Q3 <- "./../../data/processed/df_2015_Q3.RData"
input_2015_Q4 <- "./../../data/processed/df_2015_Q4.RData"
input_2016 <- "./../../data/processed/df_2016.RData"

#Splitting the test data 2016.05 (only)
input_test <- "./../../data/processed/df_test.RData"
```

## Getting the data

```{r loadingTheData, eval=FALSE}
df <- fread(input = input_file, header = T, stringsAsFactors = F)
df$fecha_dato <- lubridate::ymd(df$fecha_dato)
df$fecha_dato_year <- lubridate::year(df$fecha_dato)
df$fecha_dato_month <- lubridate::month(df$fecha_dato)

#Splitting in test & training
idx_train <- df$fecha_dato_year <= 2016 & df$fecha_dato_month <= 4
df_train <- df[idx_train,]
save(df_train, file = input_train)

#Splititng Data horizontally by time
idx_2015_Q1 <- df$fecha_dato_year == 2015 & df$fecha_dato_month <= 3
idx_2015_Q2 <- df$fecha_dato_year == 2015 & (df$fecha_dato_month > 3 & df$fecha_dato_month <= 6)
idx_2015_Q3 <- df$fecha_dato_year == 2015 & (df$fecha_dato_month > 6 & df$fecha_dato_month <= 9)
idx_2015_Q4 <- df$fecha_dato_year == 2015 & df$fecha_dato_month > 9

df_2015_Q1 <- df[idx_2015_Q1,]
df_2015_Q2 <- df[idx_2015_Q2,]
df_2015_Q3 <- df[idx_2015_Q3,]
df_2015_Q4 <- df[idx_2015_Q4,]

idx_2016_til_april <- df$fecha_dato_year == 2016 & df$fecha_dato_month <= 4
idx_test <- df$fecha_dato_year == 2016 & df$fecha_dato_month > 4
df_2016 <- df[idx_2016_til_april,]
df_test <- df[idx_test,]

save(df_2015_Q1, file = input_2015_Q1)
save(df_2015_Q2, file = input_2015_Q2)
save(df_2015_Q3, file = input_2015_Q3)
save(df_2015_Q4, file = input_2015_Q4)

save(df_2016, file = input_2016)
save(df_test, file = input_test)
rm(list = c("df", "df_2015_Q1", "df_2015_Q2", "df_2015_Q3", "df_2015_Q4", "df_2016", "df_test", "df_train"))
rm(list = ls(pattern = "idx_*"))
```

## Data Exploration

```{r readTheData, cache= T, eval = F}
load(input_train)
df_train[,fecha_dato_year:= NULL]
df_train[,fecha_dato_month := NULL]

#Training dataset
#Dimension

dim(df_train)

#Structure 
str(df_train)

#Summary
summary(df_train)
```


```{r setDataTypes, eval = F}
setDataTypes <- function(x){
    x$ncodpers <- as.factor(x$ncodpers)
    x$ind_empleado <- as.factor(x$ind_empleado)
    x$pais_residencia <- as.factor(x$pais_residencia)
    x$sexo <- as.factor(x$sexo)
    
    x$fecha_alta <- lubridate::ymd(x$fecha_alta)
    
    x$ind_nuevo <- as.factor(x$ind_nuevo)
    x$indrel <- as.factor(x$indrel)
    
    x$ult_fec_cli_1t <- lubridate::ymd(x$ult_fec_cli_1t)
    
    x$indrel_1mes <- as.factor(x$indrel_1mes)
    x$tiprel_1mes <- as.factor(x$tiprel_1mes)
    x$indresi <- as.factor(x$indresi)
    x$indext <- as.factor(x$indext)
    x$conyuemp <- as.factor(x$conyuemp)
    x$canal_entrada <- as.factor(x$canal_entrada)
    x$indfall <- as.factor(x$indfall)
    x$tipodom <- as.factor(x$tipodom)
    x$cod_prov <- as.factor(x$cod_prov)
    x$nomprov <- as.factor(x$nomprov)
    x$ind_actividad_cliente <- as.factor(x$ind_actividad_cliente)
    
    x$segmento <- as.factor(x$segmento)
    x
}

#setting the proper datatypes for the features
df_train <- setDataTypes(df_train)
save(df_train, file = input_train_typesSet)
```

```{r loadTrainWithTypesSet, cache=T}
load(input_train_typesSet)

#structure
str(df_train)

#summary
#summary(df_train)
```

### Missing Values/ Empty Values Analysis

Working with all of the data from beginning til April 2016 (included) and trying to answer the following question __Which feature has missing values? Or Which feature (either character or factor) has empty values?__

```{r missingValues}
noOfMissingValues <- function(x){
    #all features
    sum(is.na(x))
}

noOfEmptyValues <- function(x){
    #For character vector and factors
    if(is.character(x) | is.factor(x)){
     sum(x == "")   
    }else{
        0
    }
}

tbl_missing_nas <- sort(sapply(as.list(df_train), FUN = noOfMissingValues), decreasing = T)
tbl_missing_nas <- as.matrix(tbl_missing_nas[tbl_missing_nas > 0])
colnames(tbl_missing_nas) <- "NAs"

tbl_empty <- sort(sapply(as.list(df_train), FUN = noOfEmptyValues), decreasing = T)
tbl_empty <- as.matrix(tbl_empty[tbl_empty > 0])
colnames(tbl_empty) <- "EMPTYs"
```
`r kable(as.matrix(tbl_missing_nas, ncol = 1), format = "markdown", caption = "Summary OF MISSING VALUES (>0) BY FEATURE (ALL FEATURES)")`
`r kable(as.matrix(tbl_empty, ncol = 1), format = "markdown",caption = "Summary OF EMPTY VALUES (>0) BY FEATURE (CHARACTER & FACTOR ONLY)")`

__ult_fec_cli_1t__, last date as primary customer (if he isn't at the end of the month)

__renta__, Gross income of the household

```{r rentaAnalysis}
tmp <- select(df_train, fecha_dato, ncodpers, renta)
renta_an <- group_by(tmp, ncodpers) %>% 
    summarise(min = min(renta, na.rm = T), max = max(renta, na.rm = T), mean = mean(renta, na.rm = T), nas = sum(is.na(renta)))

idx_renta_rentaAvailableSomeNAs <- renta_an$nas > 0 & (!is.na(renta_an$min) | !is.na(renta_an$max) | !is.na(renta_an$mean))
idx_renta_rentaAvailableAllNAs <- renta_an$nas > 0 & (is.na(renta_an$min) & is.na(renta_an$max) & is.na(renta_an$mean))

renta_an_1 <- renta_an[idx_renta_rentaAvailableSomeNAs,]
renta_an_2 <- renta_an[idx_renta_rentaAvailableAllNAs,]
renta_an_1$sameValues <- (renta_an_1$min == renta_an_1$max) & (renta_an_1$mean == renta_an_1$max)
```

From the data below we can see that there are `r dim(renta_an_1)[1]` clients (first 10s are visible) where it is posible to recove the missing renta just just using the available renta for the same client not NA. On the other side there are `r dim(renta_an_2)[1]` clients (first 10s are visible) where such information cannot be found.

`r kable(head(renta_an_1, 10), format = "markdown",caption = "Summary OF CLIENTS (first 10s) WITH Missing renta for some entries")`
`r kable(head(renta_an_2, 10), format = "markdown",caption = "Summary OF CLIENTS (first 10s) WITH Missing renta all entries")`

```{r exampleOfMissingRentaWIthSomeValues, eval = F}
df_train[df_train$ncodpers == 183439, c(1:30), with = F]
```

```{r exampleOfMissingRentaWithAllMissing, eval = F}
df_train[df_train$ncodpers == 15917, c(1:30), with = F]
df_train[df_train$ncodpers == 16203, c(1:30), with = F]
```

__Initial strategy__ find out the median/ average salary by province using `nomprov` and `sex` as initial sefmenation and use such value for imputation of the entries with missing `renta`

```{r salaySegmentation}
salary_by_country <-  df_train %>% 
    select(pais_residencia,renta) %>%
    group_by(pais_residencia)

salary_by_country_summary <- summarise(salary_by_country, 
                            min = min(renta, na.rm = T),
                            max = min(renta, na.rm = T),
                            mean = mean(renta, na.rm = T),
                            median = median(renta, na.rm = T))

ggplot(salary_by_country_summary, aes(x = median)) +
    geom_histogram()

salary_by_country_sex <-  df_train %>% 
    select(pais_residencia,sexo, renta) %>%
    group_by(pais_residencia, sexo)

salary_by_country_sex_summary <- summarise(salary_by_country_sex, 
                            min = min(renta, na.rm = T),
                            max = min(renta, na.rm = T),
                            mean = mean(renta, na.rm = T),
                            median = median(renta, na.rm = T))

ggplot(salary_by_country_sex_summary, aes(x = median, color = sexo)) +
    geom_histogram() + facet_wrap(facet = ~sexo)

ggplot(salary_by_country_summary, aes(x = median)) +
    geom_histogram()

salary_by_location_sex <-  df_train %>% 
    select(pais_residencia,nomprov,sexo, renta) %>%
    group_by(pais_residencia,nomprov,sexo)

salary_summary <- summarise(salary_by_location_sex, 
                            min = min(renta, na.rm = T),
                            max = min(renta, na.rm = T),
                            mean = mean(renta, na.rm = T),
                            median = median(renta, na.rm = T))

```

## Tell a story

```{r theStory}
#dataset dimension
dim(df_train)

#name of the features included in the dataset
colnames((df_train))

#no of unique customers in the dataset
unique_c_ids <- unique(df_train$ncodpers)
length(unique_c_ids)

table(df_train$sexo)
sum(is.na(df_train$sexo))
sum(df_train$sexo == "")

by_sex <- df_train %>%
    select(ncodpers, sexo) %>%
    filter(sexo != "") %>%
    group_by(ncodpers, sexo) %>%
    summarise(count = n())

length(unique(by_sex$ncodpers)) == dim(by_sex)[1]

#customers with no sex info 
length(unique_c_ids) - dim(by_sex)[1]

#Sex distribution
ggplot(data = by_sex, mapping = aes(x = sexo, fill = sexo)) +
    geom_bar()
```

```{r consistencyCheckSex}
by_sex_check <- df_train %>%
    select(ncodpers, sexo) %>%
    filter(sexo != "") %>%
    group_by(ncodpers) %>%
    mutate(isH = any(sexo == "H"), isV = any(sexo == "V"))

tmp <- by_sex_check[by_sex_check$isH==T & by_sex_check$isV == T,]
```

`r kable(tmp, format = "markdown",caption = "Customers with ambiguous sex")`
