---
title: "Exploration & Data Analysis"
author: "Pier Lorenzo Paracchini"
date: "11/27/2016"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = T, warning = F, message = F)
```

Using the `tidyverse` package for performing exploration and data analysis.

```{r requiredLibraries}
require(data.table)
require(lubridate)
input_file <- "./../data/raw/train_ver2.csv"
input_2015_Q1 <- "./../data/processed/df_2015_Q1.RData"
input_2015_Q2 <- "./../data/processed/df_2015_Q2.RData"
input_2015_Q3 <- "./../data/processed/df_2015_Q3.RData"
input_2015_Q4 <- "./../data/processed/df_2015_Q4.RData"
input_2016 <- "./../data/processed/df_2016.RData"
```

## Information about the data

__Data fields__

__fecha_dato__, The table is partitioned for this column  
__ncodpers__, Customer code  
__ind_empleado__, Employee index: A active, B ex employed, F filial, N not employee, P pasive
pais_residencia	Customer's Country residence  
__sexo__, Customer's sex  
__age__, Age  
__fecha_alta__, The date in which the customer became as the first holder of a contract in the bank  
__ind_nuevo__, New customer Index. 1 if the customer registered in the last 6 months.  
__antiguedad__, Customer seniority (in months)
__indrel__	1 (First/Primary), 99 (Primary customer during the month but not at the end of the month)  
__ult_fec_cli_1t__, Last date as primary customer (if he isn't at the end of the month)  
__indrel_1mes__, Customer type at the beginning of the month ,1 (First/Primary customer), 2 (co-owner ),P (Potential),3 (former primary), 4(former co-owner)  
__tiprel_1mes__, Customer relation type at the beginning of the month, A (active), I (inactive), P (former customer),R (Potential)  

__indresi__, Residence index (S (Yes) or N (No) if the residence country is the same than the bank country)  
__indext__, Foreigner index (S (Yes) or N (No) if the customer's birth country is different than the bank country)  
__conyuemp__, Spouse index. 1 if the customer is spouse of an employee  
__canal_entrada__, channel used by the customer to join  
__indfall__, Deceased index. N/S  
__tipodom__, Addres type. 1, primary address  
__cod_prov__, Province code (customer's address)  
__nomprov__, Province name  
__ind_actividad_cliente__,Activity index (1, active customer; 0, inactive customer)  
__renta__, Gross income of the household  
__segmento__, segmentation: 01 - VIP, 02 - Individuals 03 - college graduated  

__ind_ahor_fin_ult1__, Saving Account  
__ind_aval_fin_ult1__, Guarantees  
__ind_cco_fin_ult1__, Current Accounts  
__ind_cder_fin_ult1__, Derivada Account  
__ind_cno_fin_ult1__, Payroll Account  
__ind_ctju_fin_ult1__, Junior Account  
__ind_ctma_fin_ult1__, Más particular Account  
__ind_ctop_fin_ult1__, particular Account  
__ind_ctpp_fin_ult1__, particular Plus Account  
__ind_deco_fin_ult1__, Short-term deposits  
__ind_deme_fin_ult1__, Medium-term deposits  
__ind_dela_fin_ult1__, Long-term deposits  
__ind_ecue_fin_ult1__, e-account  
__ind_fond_fin_ult1__, Funds  
__ind_hip_fin_ult1__, Mortgage  
__ind_plan_fin_ult1__, Pensions  
__ind_pres_fin_ult1__, Loans  
__ind_reca_fin_ult1__, Taxes  
__ind_tjcr_fin_ult1__, Credit Card  
__ind_valo_fin_ult1__, Securities  
__ind_viv_fin_ult1__, Home Account  
__ind_nomina_ult1__, Payroll  
__ind_nom_pens_ult1__, Pensions  
__ind_recibo_ult1__, Direct Debit  

## Getting the data

```{r loadingTheData, eval=FALSE}
df <- fread(input = input_file, header = T, stringsAsFactors = F)
df$fecha_dato <- lubridate::ymd(df$fecha_dato)
df$fecha_dato_year <- lubridate::year(df$fecha_dato)
df$fecha_dato_month <- lubridate::month(df$fecha_dato)

idx_2015_Q1 <- df$fecha_dato_year == 2015 & df$fecha_dato_month <= 3
idx_2015_Q2 <- df$fecha_dato_year == 2015 & (df$fecha_dato_month > 3 & df$fecha_dato_month <= 6)
idx_2015_Q3 <- df$fecha_dato_year == 2015 & (df$fecha_dato_month > 6 & df$fecha_dato_month <= 9)
idx_2015_Q4 <- df$fecha_dato_year == 2015 & df$fecha_dato_month > 9

df_2015_Q1 <- df[idx_2015_Q1,]
df_2015_Q2 <- df[idx_2015_Q2,]
df_2015_Q3 <- df[idx_2015_Q3,]
df_2015_Q4 <- df[idx_2015_Q4,]

df_2016 <- df[df$fecha_dato_year == 2016,]

save(df_2015_Q1, file = input_2015_Q1)
save(df_2015_Q2, file = input_2015_Q2)
save(df_2015_Q3, file = input_2015_Q3)
save(df_2015_Q4, file = input_2015_Q4)

save(df_2016, file = input_2016)
rm(list = c("df", "df_2015_Q1", "df_2015_Q2", "df_2015_Q3", "df_2015_Q4", "df_2016"))
```

## Data Exploration

```{r readTheData}
setDataTypes <- function(x){
    x <- df_2015_Q1
    x$ncodpers <- as.factor(x$ncodpers)
    x$ind_empleado <- as.factor(x$ind_empleado)
    x$pais_residencia <- as.factor(x$pais_residencia)
    x$sexo <- as.factor(x$sexo)
    x$fecha_alta <- lubridate::ymd(x$fecha_alta)
    x$ind_nuevo <- factor(x$ind_nuevo, levels = c(0,1), labels = c("NO", "YES"))
    x$indrel <- as.factor(x$indrel)
    
    idx_isEmpty <- x$ult_fec_cli_1t == ""
    x$ult_fec_cli_1t[idx_isEmpty] <- NA
    x$ult_fec_cli_1t[!idx_isEmpty] <-lubridate::ymd(x$ult_fec_cli_1t[!idx_isEmpty])
    
}
load(input_2015_Q1)
table(df_2015_Q1$fecha_dato_year, df_2015_Q1$fecha_dato_month)
```

## TODO

* github - push data
* Segmentation - vertical part & splitting in Train/ test by time
* 
