---
title: "Exploration & Data Analysis"
author: "Pier Lorenzo Paracchini"
date: "11/27/2016"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = T, warning = F, message = F)
```

Using the `tidyverse` package for performing exploration and data analysis.

```{r requiredLibraries}
require(data.table)
require(lubridate)
require(dplyr)
require(knitr)
input_file <- "./../../data/raw/train_ver2.csv"

#Splitting the training data as one file till 2016.04 (included)
input_train <- "./../../data/processed/df_train.RData"
input_train_typesSet <- "./../../data/processed/df_train_typesSet.RData"

#Splitting up the data by periods to have smaller chunks
#2015 split in qurters
#2016 from Jan to April (included)
input_2015_Q1 <- "./../../data/processed/df_2015_Q1.RData"
input_2015_Q2 <- "./../../data/processed/df_2015_Q2.RData"
input_2015_Q3 <- "./../../data/processed/df_2015_Q3.RData"
input_2015_Q4 <- "./../../data/processed/df_2015_Q4.RData"
input_2016 <- "./../../data/processed/df_2016.RData"

#Splitting the test data 2016.05 (only)
input_test <- "./../../data/processed/df_test.RData"
```

## Getting the data

```{r loadingTheData, eval=FALSE}
df <- fread(input = input_file, header = T, stringsAsFactors = F)
df$fecha_dato <- lubridate::ymd(df$fecha_dato)
df$fecha_dato_year <- lubridate::year(df$fecha_dato)
df$fecha_dato_month <- lubridate::month(df$fecha_dato)

#Splitting in test & training
idx_train <- df$fecha_dato_year <= 2016 & df$fecha_dato_month <= 4
df_train <- df[idx_train,]
save(df_train, file = input_train)

#Splititng Data horizontally by time
idx_2015_Q1 <- df$fecha_dato_year == 2015 & df$fecha_dato_month <= 3
idx_2015_Q2 <- df$fecha_dato_year == 2015 & (df$fecha_dato_month > 3 & df$fecha_dato_month <= 6)
idx_2015_Q3 <- df$fecha_dato_year == 2015 & (df$fecha_dato_month > 6 & df$fecha_dato_month <= 9)
idx_2015_Q4 <- df$fecha_dato_year == 2015 & df$fecha_dato_month > 9

df_2015_Q1 <- df[idx_2015_Q1,]
df_2015_Q2 <- df[idx_2015_Q2,]
df_2015_Q3 <- df[idx_2015_Q3,]
df_2015_Q4 <- df[idx_2015_Q4,]

idx_2016_til_april <- df$fecha_dato_year == 2016 & df$fecha_dato_month <= 4
idx_test <- df$fecha_dato_year == 2016 & df$fecha_dato_month > 4
df_2016 <- df[idx_2016_til_april,]
df_test <- df[idx_test,]

save(df_2015_Q1, file = input_2015_Q1)
save(df_2015_Q2, file = input_2015_Q2)
save(df_2015_Q3, file = input_2015_Q3)
save(df_2015_Q4, file = input_2015_Q4)

save(df_2016, file = input_2016)
save(df_test, file = input_test)
rm(list = c("df", "df_2015_Q1", "df_2015_Q2", "df_2015_Q3", "df_2015_Q4", "df_2016", "df_test", "df_train"))
rm(list = ls(pattern = "idx_*"))
```

## Data Exploration

```{r readTheData, cache= T, eval = F}
load(input_train)
df_train[,fecha_dato_year:= NULL]
df_train[,fecha_dato_month := NULL]

#Training dataset
#Dimension

dim(df_train)

#Structure 
str(df_train)

#Summary
summary(df_train)
```


```{r setDataTypes, eval = F}
setDataTypes <- function(x){
    x$ncodpers <- as.factor(x$ncodpers)
    x$ind_empleado <- as.factor(x$ind_empleado)
    x$pais_residencia <- as.factor(x$pais_residencia)
    x$sexo <- as.factor(x$sexo)
    
    x$fecha_alta <- lubridate::ymd(x$fecha_alta)
    
    x$ind_nuevo <- as.factor(x$ind_nuevo)
    x$indrel <- as.factor(x$indrel)
    
    x$ult_fec_cli_1t <- lubridate::ymd(x$ult_fec_cli_1t)
    
    x$indrel_1mes <- as.factor(x$indrel_1mes)
    x$tiprel_1mes <- as.factor(x$tiprel_1mes)
    x$indresi <- as.factor(x$indresi)
    x$indext <- as.factor(x$indext)
    x$conyuemp <- as.factor(x$conyuemp)
    x$canal_entrada <- as.factor(x$canal_entrada)
    x$indfall <- as.factor(x$indfall)
    x$tipodom <- as.factor(x$tipodom)
    x$cod_prov <- as.factor(x$cod_prov)
    x$nomprov <- as.factor(x$nomprov)
    x$ind_actividad_cliente <- as.factor(x$ind_actividad_cliente)
    
    x$segmento <- as.factor(x$segmento)
    x
}

#setting the proper datatypes for the features
df_train <- setDataTypes(df_train)
save(df_train, file = input_train_typesSet)
```

```{r loadTrainWithTypesSet, cache=T}
load(input_train_typesSet)

#structure
str(df_train)

#summary
summary(df_train)
```

### Missing Values/ Empty Values Analysis

Working with all of the data from beginning til April 2016 (included) and trying to answer the following question __Which feature has missing values? Or Which feature (either character or factor) has empty values?__

```{r missingValues}
noOfMissingValues <- function(x){
    #all features
    sum(is.na(x))
}

noOfEmptyValues <- function(x){
    #For character vector and factors
    if(is.character(x) | is.factor(x)){
     sum(x == "")   
    }else{
        0
    }
}

tbl_missing_nas <- sort(sapply(as.list(df_train), FUN = noOfMissingValues), decreasing = T)
tbl_missing_nas <- as.matrix(tbl_missing_nas[tbl_missing_nas > 0])
colnames(tbl_missing_nas) <- "NAs"

tbl_empty <- sort(sapply(as.list(df_train), FUN = noOfEmptyValues), decreasing = T)
tbl_empty <- as.matrix(tbl_empty[tbl_empty > 0])
colnames(tbl_empty) <- "EMPTYs"
```
`r kable(as.matrix(tbl_missing_nas, ncol = 1), caption = "Summary OF MISSING VALUES (>0) BY FEATURE (ALL FEATURES)")`
`r kable(as.matrix(tbl_empty, ncol = 1), caption = "Summary OF EMPTY VALUES (>0) BY FEATURE (CHARACTER & FACTOR ONLY)")`

__ult_fec_cli_1t__, last date as primary customer (if he isn't at the end of the month)

__renta__, Gross income of the household

```{r rentaAnalysis}
tmp <- select(df_train, fecha_dato, ncodpers, renta)
renta_an <- group_by(tmp, ncodpers) %>% 
    summarise(min = min(renta, na.rm = T), max = max(renta, na.rm = T), mean = mean(renta, na.rm = T), nas = sum(is.na(renta)))

idx_renta_rentaAvailableSomeNAs <- renta_an$nas > 0 & (!is.na(renta_an$min) | !is.na(renta_an$max) | !is.na(renta_an$mean))
idx_renta_rentaAvailableAllNAs <- renta_an$nas > 0 & (is.na(renta_an$min) & is.na(renta_an$max) & is.na(renta_an$mean))

renta_an_1 <- renta_an[idx_renta_rentaAvailableSomeNAs,]
renta_an_2 <- renta_an[idx_renta_rentaAvailableAllNAs,]
renta_an_1$sameValues <- (renta_an_1$min == renta_an_1$max) & (renta_an_1$mean == renta_an_1$max)
```

From the data below we can see that there are `r dim(renta_an_1)[1]` clients where it is posible to recove the missing renta just just using the available renta for the same client not NA. On teh other side there are `r dim(renta_an_2)[1]` clients where such information cannot be found.

`r kable(renta_an_1, caption = "Summary OF CLIENTS WITH Missing renta for some entries")`
`r kable(head(renta_an_2, 20), caption = "Summary OF CLIENTS (first 20s) WITH Missing renta all entries")`
